---
title: "IBD Patient Data Practice"
author: "Peter Higgins"
date: "9/15/2018"
output: html_document
---

# IBD Patient Growth

## Setting up
1. first open a New Project called mi_maps_2018
2. check that there is a corresponding folder on your hard drive
3. Save files to this folder, including "ibd_2018_tidy.xlsx"  and "mi_cty_pop.xlsx"
4. Confirm that these files now appear under the Files tab in your RStudio project
5. install the pacman package (see first chunk below) if you don't already have this one.
6. Run the chunks below and follow the instructions
7. use themes and YAML options to add style to your HTML document, including floating table of contents and numbered sections

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("pacman")  
#Only if you don't already have pacman
pacman::p_load(tidyverse, magrittr, rmarkdown, knitr, readxl, acs, choroplethr, choroplethrMaps, kableExtra)
```

## Data Import

Use the chunk below to bring in 2018 data from the file
"ibd_2018_tidy.xlsx" 


```{r bring_in_data}
# import code here
ibd_2018  <- read_excel("ibd_2018_tidy.xlsx") 
```



# Initial Plot

Make a plot and describe it here

Instructions:
It should be a line plot of the top **10** counties by IBD patients seen in 2018 at Michigan Medicine.
x axis should include 2016-2018
Unique color in legend for each county

```{r plot1}
# code here for line graph
library(ggplot2)
install.packages("forcats")
library(forcats)
ibd_plot <- ibd_2018 %>%
  filter(year == "2018") %>%
top_n(count, n=10) %>%
arrange(desc(count))
ggplot(ibd_2018, aes(x = county, y = count)) +
    geom_line(aes(linetype= "dashed")) + #,color = as.factor(county))) 
    geom_point() +
    labs(
      x = "Michigan counties",
      y = "Count",
      title = "Title",
      subtitle = "Subtitle",
      caption = "Caption"
    )
```


# How to Table

You can also make tables, for example:
Look up some formatting options with kableExtra at:
<https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html>

```{r table1, echo=FALSE}
mi_pop  <- read_excel("mi_cty_pop.xlsx") 
mi_pop %>% 
  top_n(pop, n=3) %>% 
  arrange(desc(pop)) %>% 
  select(county, pop) %>% 
  kable() %>% 
  kable_styling()
```


# A Second Table

You can also embed tables, for example:
use knitr::kable and kableExtra package to make a nicely formatted table of the counties with the top 20 number of IBD patients seen at Michigan Medicine in 2018, arranged in descending order.
Use some kableExtra formatting to make it look good.

```{r table2, echo=FALSE}
install.packages("kableExtra")
library(knitr)
library(kableExtra)

ibd_2018 %>%
  top_n(count, n=20) %>%
arrange(desc(count)) %>% 
  select(county, count) %>% 
  kable() %>% 
  kable_styling()
```

# Challenge Graph

Use your skills to
1. Estimate number of IBD patients who are NOT coming in each of top **10** counties based on county population
2. Use estimator of 50 IBD patients per thousand population from our neighbors in Ontario to calculate how many IBD patients there should be
3. Line Plot the number of patients in each of top 10 counties NOT coming to UM (y axis) with year on the x axis, only 2016-18

Note that you may want to use the tolower() function to make your left_join work.

```{r graph2, echo=FALSE}
# code here
```

# Use Rmarkdown formatting 
write up your conclusions, and proposals about where to market the IBD program - which counties have the most IBD patients that are not being seen at Michigan Medicine within a reasonable distance?

** remember to**
use themes and YAML options to add style to your HTML document, including floating table of contents and numbered sections